\documentclass[journal, a4paper]{IEEEtran}

\include{packages}

\newcommand{\reffig}[1]{Fig. \ref{#1}}

\hyphenation{op-tical net-works semi-conduc-tor}


\begin{document}

\title{Survey paper of techniques for optimizing DRAM refresh}

\author{\IEEEauthorblockN{Dan Larsson, Jonas Hemlin}\\
\IEEEauthorblockA{Master Program of Embedded Electronic Systems Design\\
Chalmers University of Technology, Gothenburg, Sweden\\
Email: \{larsdan, jhemlin\}@student.chalmers.se}}

\maketitle


\begin{abstract}
We survey nine different techniques for lowering the number of refresh operations performed in a modern DRAM system. Because the portion of power consumption coming from refresh operations in DRAM is high and increasing, these techniques are both topical and highly considerable. The relative performance of the techniques is compared, and some projections for the future concerning higher density and bigger capacity DRAMs are discussed. The new ways to improve DRAM, like 3D stacked DRAM, will increase the need to mitigate refresh power even further. To give a context to the techniques surveyed the basic operation of DRAM is elaborated on. 

All the techniques surveyed fits into one or more approach, where each approach utilizes on of four types of information, Retention Time, Access Recency, Data Tolerance, or Validity. Among these approaches, we believe that Retention Time and Validity are the most promising as their benefits manage to scale well with DRAM capacity. Moreover, they show large improvements to a relatively low complexity, compared to the others.
\end{abstract}

\input{include/introduction}
\input{include/background}
\input{include/reducing_refresh_power}
\input{include/discussion}
\input{include/summary}

\section*{Acknowledgment}
The authors would like to thank Sally A. McKee for her guidance during the creation of this survey.



% \IEEEtriggercmd{\enlargethispage{-5.5in}}
\IEEEtriggercmd{\newpage}
\IEEEtriggeratref{12}


\bibliographystyle{../IEEEtran}
\bibliography{../IEEEfull,../references}

\end{document}


